SunPy - Python for Solar Physics

### 2014-01-28 1-Introduction.tex

The design philosophy of SunPy is to provide a clean, simple-to-use, and well 
structured package that provides the \textit{core} tools for solar-physics data analysis. The 
primary focus of SunPy's early development is to provide specialised, linked, 
data types that allow the acquisition, processing and visualisation of all types 
of solar data.

The purpose of this paper is to provide an overview of SunPy's current 
capabilities, an overview of the development model and community aspects of the 
SunPy project, as well as future plans. The latest release of SunPy (0.4)
can be downloaded from
the official website (\url{http://sunpy.org}) or can be installed using 
the Python package index (\mbox{\url{http://pypi.python.org/pypi}}).
%schriste - the mbox prevents latex from breaking the url across different lines.

>>>>

The design philosophy of SunPy is to be a clean, simple-to-use, and well 
structured package that provides the \textit{core} tools for solar-physics data 
analysis. The primary focus of SunPy's early development is to provide 
specialised, linked, data types that allow the acquisition, processing and 
visualisation of all types of solar data.

The purpose of this paper is to provide an overview of SunPy's current 
capabilities, an overview of the development model and community aspects of the 
SunPy project, as well as future plans. The latest release of SunPy (0.4)
can be downloaded from the \href{http://sunpy.org}{official website} or can be
installed using the \href{http://pypi.python.org/pypi}{Python package index}.

### 2014-01-28 2-Data_Types.tex

The primary focus of SunPy is to provide data structures 
that are specifically designed for the primary varieties of solar physics data, namely
images, time series, and spectra. SunPy provides three core data
types using Python classes:
\texttt{Map} (2D spatial data),
\texttt{LightCurve} (1D temporal series)
and \texttt{Spectrum}/\texttt{Spectrogram} (1 and 2D spectra).

The classes provide access to the original data
along with associated metadata and provide appropriate convenience functions to
enable data analysis and visualisation. For each of these classes, the data is
stored in the \texttt{.data} attribute while the metadata is stored 
in the \texttt{.meta} attribute. It is 
possible to instantiate the data types from various different 
sources: \textit{e.g.}, files, URLs, arrays or time ranges.  In order to provide
instrument-specific specialisation, all the core SunPy data types support specialisation;
\textit{e.g.}, \texttt{Map} has an \texttt{AIAMap} sub-type. 
Data visualisation is provided by two functions: \texttt{peek()}, for quick plotting,
and \texttt{plot()}, for more fine-grained control plotting that integrates with 
\texttt{matplotlib}.
% LightCurve does NOT have a meta attribute.  Needs to be fixed in the
% re-write of LightCurve

@@

\subsection{Visualisation}
\label{subsec:Viz}
As has been demonstrated in this section, the core SunPy datatypes 
include visualisation code that is tailored to that data type. 
These visualisation methods all currently utilise the matplotlib 
package, and are designed in such a way that they integrate well with 
the \texttt{pyplot} functional interface of \texttt{matplotlib}, as demonstrated in 
Listing \ref{code:mpl_1}.

%schriste - I vote for removing all of the following code as being too
%down in the weeds
%ji -  I agree.  Do we even need to say much more than give a brief
%outline of the .peek and .plot methods?
\begin{listing}[H]
\begin{minipage}{0.49\columnwidth}
\begin{minted}[bgcolor=bg]{python}
import sunpy.map
import matplotlib.pyplot as plt

#Read in a aia file using SunPy
aiamap = SunPy.map.Map('aia.fits')

#Create the figure and axis 
#explicitly
fig = plt.figure()
ax = plt.subplot()

#Create a image plot with raw data
im = ax.imshow(aiamap.data,...)
#add a colour bar
plt.colorbar(im,ax=ax)

#Add a title
plt.title("My AIA plot")
plt.show()
\end{minted}
\end{minipage}
\begin{minipage}{0.49\columnwidth}
\begin{minted}[bgcolor=bg]{python}
import SunPy.map
import matplotlib.pyplot as plt

#Read in a aia file using SunPy
aiamap = SunPy.map.Map('aia.fits')

#Create the figure and axis 
#explicitly
fig = plt.figure()
ax = plt.subplot()

#Plot the AIA map
im = aiamap.plot(axes=ax)
#add a colour bar
plt.colorbar(im,ax=ax)

#Add a title
plt.title("My AIA plot")
plt.show()
\end{minted}
\end{minipage}
\caption{A simple example of how SunPy's plotting functions provide similar 
behaviour to the \texttt{matplotlib.pyplot} interface.}
\label{code:mpl_1}
\end{listing}

@@

%ayshih - I recommend removing the following sentences
More technically what is done in the plotting routines is firstly, to 
make sure that the current state of the figure and axes is maintained 
in the same way as it is inside the \texttt{pyplot} functions. For 
\texttt{Map} this means setting \texttt{pyplot.sci()} or set current 
image, so that \texttt{plt.colorbar()} 
automatically detects the plotted \texttt{AxesImage} object. 
Secondly, it also means returning the same type of object as the 
\texttt{pyplot} interfaces function, \textit{i.e.} for \texttt{Map} 
it returns the \texttt{AxesImage} instance which is the same type of 
return as \texttt{pyplot.imshow}.

>>>>

Currently SunPy provides data structures that are specifically designed for the
three primary varieties of solar physics data, namely images, time series, and
spectra. SunPy provides core data types for these using Python classes:
\texttt{Map} (2D spatial data), \texttt{LightCurve} (1D temporal series)
and \texttt{Spectrum}/\texttt{Spectrogram} (1 and 2D spectra).

The classes provide access to the original data
along with associated metadata and provide appropriate convenience functions to
enable data analysis and visualisation. For each of these classes, the data is
stored in the \texttt{.data} attribute while the metadata is stored 
in the \texttt{.meta}\footnote{Except \texttt{LightCurve} which currently does 
not.} attribute. It is possible to instantiate the data types from various 
different sources: \textit{e.g.}, files, URLs, arrays or time ranges.  In order 
to provide instrument-specific specialisation, all the core SunPy data types 
support subclassing; \textit{e.g.}, \texttt{Map} has an \texttt{AIAMap} 
sub-type. 
Data visualisation is provided by two functions: \texttt{peek()}, for quick 
plotting, and \texttt{plot()}, for plotting with more fine-grained control that 
integrates with \texttt{matplotlib}.

@@

\subsection{Visualisation}
\label{subsec:Viz}
As has been demonstrated in this section, the core SunPy datatypes 
include visualisation code that is tailored to that data type. 
These visualisation methods all currently utilise the \texttt{matplotlib} 
package, and are designed in such a way that they integrate well with 
the \texttt{pyplot} functional interface of \texttt{matplotlib}.

@@

### 2014-01-28 3-1-VSO.tex

The Virtual Solar Observatory (VSO) provides a single, standard query
interface to access solar data from different archives around the world
 \citep{hill2009}.
Data products can be requested for specific instruments or missions, but
can also be requested based on physical parameters of the data product such
as the wavelength range.
In addition to the VSO's primary, web-based interface
(\url{http://virtualsolar.org}), a SOAP (Simple Object Access Protocol)
service is also available.
SunPy's \texttt{vso} module accesses the VSO via the SOAP service using the
\texttt{suds} package.
This module was developed under the European Space Agency's Summer of Code in 
Space (ESA's SOCIS) of 2011.

Listing~\ref{code:vso_query_simple} shows an example of how to query the VSO
using the \texttt{vso} module.
Each query is constructed by one or more attributes objects, where each
attribute object is a constraint on a parameter of the data set, such as the
time, instrument, wavelength, or  other properties.
It also shows how to download the data using the result
from the query. One can use the metadata associated with the data sets to specify how the
files will be named locally, such as creating subdirectories based on the
instrument names.

Listing~\ref{code:vso_query_advanced} shows examples of how to make advanced
queries by combining attribute objects.
Two attribute objects can be combined with a logical \texttt{or} operation
using the ``$|$'' operator.
All attribute objects provided to the query as keyword arguments are combined
with a logical \texttt{and} operation.

@@

%schriste - what about just listing the sources as an array to instead of doing a show
%           to make the point that the | query worked appropriately? That would make this
%            listing much shorter.
\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
#Querying multiple instruments/missions over the same time range
>>> instrument_condition = ( vso.attrs.Instrument('lasco') |
...                          vso.attrs.Detector('cor1') )
>>> coronagraphs = client.query(vso.attrs.Time(tstart, tend),
...                             instrument_condition)
>>> coronagraphs.num_records()
485
>>> coronagraphs.show()
Start time           End time             Source    Instrument  Type
----------           --------             ------    ----------  ----
2011-06-07 05:35:23  2011-06-07 05:35:48  SOHO      LASCO       CORONA
...
2011-06-07 10:25:43  2011-06-07 10:25:45  STEREO_B  SECCHI      CORONA
2011-06-07 10:30:00  2011-06-07 10:30:01  STEREO_A  SECCHI      CORONA

>>>>

The Virtual Solar Observatory (\href{http://virtualsolar.org}{VSO}) provides a 
single, standard query interface to access solar data from different archives 
around the world \citep{hill2009}.
Data products can be requested for specific instruments or missions, but
can also be requested based on physical parameters of the data product such
as the wavelength range.
In addition to the VSO's primary, web-based interface, a SOAP (Simple Object 
Access Protocol) service is also available.
SunPy's \texttt{vso} module accesses the VSO via the SOAP service using the
\texttt{suds} package.
This module was developed under the European Space Agency's Summer of Code in 
Space (ESA's SOCIS) of 2011.

Listing~\ref{code:vso_query_simple} shows an example of how to query the VSO
using the \texttt{vso} module.
Each query is constructed by one or more attribute objects, where each
attribute object is a constraint on a parameter of the data set, such as the
time, instrument, wavelength, or  other properties.
It also shows how to download the data using the result
from the query. One can use the metadata associated with the data sets to 
specify how the files will be named locally, such as creating subdirectories 
based on the instrument names.

Listing~\ref{code:vso_query_advanced} shows examples of how to make advanced
queries by combining attribute objects.
Two attribute objects can be combined with a logical \texttt{or} operation
using the ``$|$'' operator.
All attribute objects provided to the query as arguments are combined with a 
logical \texttt{and} operation.

@@

%schriste - what about just listing the sources as an array to instead of doing a show
%           to make the point that the | query worked appropriately? That would make this
%            listing much shorter.
%Stuart: I suggest removing the second query from here, it dosen't add much.
\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
#Querying multiple instruments/missions over the same time range
>>> instrument_condition = ( vso.attrs.Instrument('lasco') |
...                          vso.attrs.Detector('cor1') )
>>> coronagraphs = client.query(vso.attrs.Time(tstart, tend),
...                             instrument_condition)
>>> coronagraphs.num_records()
485
>>> coronagraphs.show()
Start time           End time             Source    Instrument  Type
----------           --------             ------    ----------  ----
2011-06-07 05:35:23  2011-06-07 05:35:48  SOHO      LASCO       CORONA
...
2011-06-07 10:25:43  2011-06-07 10:25:45  STEREO_B  SECCHI      CORONA
2011-06-07 10:30:00  2011-06-07 10:30:01  STEREO_A  SECCHI      CORONA

### 2014-01-28 3-4-Helioviewer.tex

SunPy provides the ability to download images hosted by the
Helioviewer Project (\url{http://helioviewer.org}).  
The aim of the Helioviewer Project is to enable
the exploration of solar and heliospheric data from multiple data
sources (such as instrumentation and feature/event catalogues) via
easy-to-use visual interfaces.  The primary data provided by the
Helioviewer Project are images in the JPEG2000 format (for further
details on the Helioviewer Project see \cite{muller2009}). The
JPEG2000 files are typically highly compressed compared to the source
FITS files they are generated from, and so can be a used as a shortcut
to visualise large amounts of data from multiple sources.  SunPy is
also used in Helioviewer production servers to manage the download and
ingestion of JPEG2000 files from remote servers.

@@

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy.net.helioviewer import HelioviewerClient
>>> hv = HelioviewerClient()
>>> hv.download_png('2099/01/01', 4.8, "[SDO,AIA,AIA,304,1,100]",
...                 x0=0, y0=0, width=512, height=512)
\end{minted}
\includegraphics[width=0.8\columnwidth]{helioviewer_latest_aia_304}
%centered image?
\caption{Acquisition of a PNG file showing the latest AIA 304\AA\ image 
available at \url{www.helioviewer.org}.}
\label{code:hv:downloadlatestpng}
\end{listing}

@@

It should be noted that the filename of the returned file has the date
and time of the requested time, which is not the time of the
observation shown in the PNG image.  This is because of the
Helioviewer API.  The API request finds images closest to the
requested time. But since the user may ask for images from multiple
sources, and each of them may have a different observation time, it is
unclear which time is the most appropriate to associate with the
resultant image.  The Helioviewer API does not select between any of
the image observation times, but instead returns an image filename
based on the request time, resulting in predictable filenames.

@@

Reading JPEG2000 file into a SunPy session requires installing two
other pieces of software. The first, \texttt{OpenJPEG}, is an open
source library for reading and writing JPEG2000 files
(\url{http://www.openjpeg.org}).  The other package required is
\texttt{glymur}, (\url{https://github.com/quintusdias/glymur}), an
interface between Python and the OpenJPEG libraries (note that these
packages are {\it not} required to use the functionality described in
Section \ref{sssec:hv:png}).

Listing \ref{code:downloadjp2} demonstrates the querying, downloading,
reading and conversion of a Helioviewer JPEG2000 file into a SunPy Map
object.  This functionality allows users to visualise and manipulate
Helioviewer-supplied image data in an identical fashion to a SunPy map
object generated from FITS data (see Section \ref{sec:map}).

>>>>

SunPy provides the ability to download images hosted by the
\href{http://helioviewer.org}{Helioviewer Project}.  
The aim of the Helioviewer Project is to enable
the exploration of solar and heliospheric data from multiple data
sources (such as instrumentation and feature/event catalogues) via
easy-to-use visual interfaces.  The primary data provided by the
Helioviewer Project are images in the JPEG2000 format (for further
details on the Helioviewer Project see \cite{muller2009}). The
JPEG2000 files are typically highly compressed compared to the source
FITS files they are generated from, and so can be a used as a shortcut
to visualise large amounts of data from multiple sources.  SunPy is
also used in Helioviewer production servers to manage the download and
ingestion of JPEG2000 files from remote servers.

@@

%Stuart: Do we need this listing?? The same points can be made with 14
\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy.net.helioviewer import HelioviewerClient
>>> hv = HelioviewerClient()
>>> hv.download_png('2099/01/01', 4.8, "[SDO,AIA,AIA,304,1,100]",
...                 x0=0, y0=0, width=512, height=512)
\end{minted}
\centering\includegraphics[width=0.5\columnwidth]{helioviewer_latest_aia_304}
\caption{Acquisition of a PNG file showing the latest AIA 304\AA\ image 
available at \url{www.helioviewer.org}.}
\label{code:hv:downloadlatestpng}
\end{listing}

@@

It should be noted that the filename of the returned file has the date
and time of the requested time, which is not the time of the observation shown 
in the PNG image.  This is because the Helioviewer API request finds images 
closest to the requested time. However, since the user may ask for images from 
multiple sources, and each of them may have a different observation time, it is
unclear which time is the most appropriate to associate with the resultant 
image.  The Helioviewer API therefore does not select between any of the image 
observation times, but instead returns an image filename based on the request 
time, resulting in predictable filenames.

@@

Reading JPEG2000 file into a SunPy session requires installing two
other pieces of software. The first, 
\href{http://www.openjpeg.org}{\texttt{OpenJPEG}}, is an open
source library for reading and writing JPEG2000 files.  The other package 
required is 
\href{https://github.com/quintusdias/glymur}{\texttt{glymur}}, an
interface between Python and the OpenJPEG libraries (note that these
packages are {\it not} required to use the functionality described in
Section \ref{sssec:hv:png}).

Listing \ref{code:downloadjp2} demonstrates the querying, downloading,
reading and conversion of a Helioviewer JPEG2000 file into a SunPy Map
object.  This functionality allows users to visualise and manipulate
Helioviewer-supplied image data in an identical fashion to a SunPy map
object generated from FITS data (see Section \ref{ssec:map}).

### 2014-01-28 authors.tex

\author{
% People who have directly contributed to the text, still in no order.
Stuart J Mumford$^1$,
Steven Christe$^2$,
David P\'erez-Su\'arez$^3$,
Andrew R Inglis$^{2,5}$,
Jack Ireland$^{6,2}$,
Albert Y Shih$^2$,
% Contributors to SunPy Sorted thus:
% No. Commits to main repo + commits in open PR's
% Date of most recent contribution
F Mayer, %721
K Hughitt, %446
R Hewett, %88
Nabil Freij$^1$, %52
M Earnshaw, %52
T Meszaros, %17
N Chorley, %17
Samuel M Bennett$^1$, %14
M Malocha, %10
J Evans, %10
A Maverick, %8
M Bates, %6
Andrew J Leonard$^7$, %3
J Ballew, %3
T Robitaille, %2
B Mampaey, %2
%Hypnus1803 %1
and 
A Kustov %1
}

@@

$^7$Department of Mathematics and Physics, Aberystwyth University, 
Physical Sciences Building, Aberystwyth, Ceredigion, SY23 3BZ,
}

>>>>

\author{
% People who have directly contributed to the text, still in no order.
Stuart J Mumford$^1$,
Steven Christe$^2$,
David P\'erez-Su\'arez$^3$,
Andrew R Inglis$^{2,5}$,
Jack Ireland$^{6,2}$,
Albert Y Shih$^2$,
Simon Liedtke$^11$,
Russell J Hewett$^8$, %88
% Contributors to SunPy Sorted thus:
% No. Commits to main repo + commits in open PR's
% Date of most recent contribution
F Mayer, %721
K Hughitt, %446
Nabil Freij$^1$, %52
M Earnshaw, %52
Tomas Meszaros$^{10}$, %17
N Chorley, %17
Samuel M Bennett$^1$, %14
M Malocha, %10
J Evans, %10
A Maverick, %8
M Bates, %6
Andrew J Leonard$^7$, %3
J Ballew, %3
Thomas P Robitaille, %2
B Mampaey, %2
%Hypnus1803 %1
and 
A Kustov %1
}

@@

$^7$Department of Mathematics and Physics, Aberystwyth University, 
Physical Sciences Building, Aberystwyth, Ceredigion, SY23 3BZ,

$^8$Department of Mathematics, Massachusetts Institute of Technology, 77 
Massachusetts Ave, E17-317, Cambridge, MA 02142,

$^9$ Max-Planck-Institut f\"{u}r Astronomie, K\"onigstuhl 17, Heidelberg
69117, Germany,

$^{10}$ Masaryk University, Faculty of Informatics, Botanicka 68a Brno,

$^{11}$ University of Bremen, Bibliothekstra\ss e 1, 28359 Bremen, Germany,
}

### 2014-01-28 solarsoft.tex

\section{Comparison to SolarSoft}
This section compares the SunPy library that is described in Sections 
\ref{sec:DataTypes}--\ref{sec:util} with the current de-facto standard solar 
physics data analysis library: SolarSoft (SSW) \citep{freeland1998}.
SolarSoft is a set of integrated software libraries, databases, and system 
utilities which provide a common programming and data-analysis environment for 
solar physics. SSW is modular in nature, which allows users to add required 
packages which contain specialised software for different instrument or tools.

SSW relies upon IDL (Interactive Data 
Language), a commercial and closed-source interactive data analysis environment 
sold by Exelis, a global aerospace, defence, information and services company. 
The SSW `gen' package provides the base analysis capabilities such as time-series
analysis, time conversions, spectral fitting, image display, and file 
I/O. It is composed of 89\% IDL code while the remainder is mostly Perl (5.4\%) 
and csh scripts (3.8\%). 
According to data generated using David A. Wheeler's `SLOCCount'\footnote{\url{http://www.dwheeler.com/sloccount/}}, SSW/gen is composed of 
289,724 lines of code with a total approximate value of \$10.4M (based on the 
Basic COCOMO model \citep{_cocomo_2014}). 
%stuart - This data was calculated using...} If we can add them to the repository 
%will improve the quality of the paper and we can always repeat the calculation in the future.
%schriste - i don't know what you mean by this 
Individual missions and instruments provide additional optional packages to 
support and process their data. For example, the SDO package \citep{sdo} is 
100\% IDL code and is 11k lines of code (\$360k) and the SOHO package 
\citep{soho} is 100k lines of code (\$3M) and is 99.9\% IDL code. SSW is hosted 
and distributed by the Lockheed Martin Solar and Astrophysics Laboratory.
%The RHESSI package is 99.97\% IDL code and is 145k total lines of code (\$5M)
While SSW itself is open-source and freely available, the IDL core is not. The cost 
of IDL is prohibitive for many undergraduates as well as institutions 
with limited means.
Even though SSW's source is open, the development of the source is primarily closed,
with no open issue tracker or established mechanism for community contributions.

One of SunPy's key aims is to provide a free and modern alternative to the 
SolarSoft (gen) library. This is made possible by the recent rise of the 
scientific Python community. Python, as described in Section \ref{sec:Intro}, 
is a general purpose high-level programming language which is free to use and 
cross-platform and powerful programming language.
It therefore has a very large user community which extends far outside of the 
solar physics or astronomy communities. It is very extensible with C, C++, 
FORTRAN or even IDL. Many books and on-line documentation is available. For 
example, at the time writing a search 
%
%stuart - \footnote{We need to be careful with this! IDL means more things that our 
%beloved Interactive Data Language. If we search for that and programming 
%then we get just 7 books! - from which 5 are just the same. I don't know how we 
%could make a more accurate search.
%If we script it it would be amazing!  Also, I would just provide the order
%of magnitude, and not a exact number, i.e. 190 and 1800 results}
% 
% schriste - I think we defined IDL properly. I was being generous with the book search for IDL
on Amazon for books on IDL programming generated $189$ results while a search 
for Python yielded $1\,788$ results. Python has a strong emphasis on 
readability a concept which is well-aligned with the scientific endeavour.
Additionally, Python is increasingly being taught in universities, which means 
that research projects can begin quickly without the need for students to learn 
a new (and relatively niche) language. Finally, knowledge of Python is a skill 
which is useful and well compensated outside of the scientific community 
(average salary of \$110k in the Washington, DC, area according to indeed.com at 
the time of writing). Python continues to see increased use in the astronomy 
community \citep{greenfield2011} which has similar goals and requirements as 
the solar physics community. Finally, Python integrates well with many 
technologies such as web servers \citep{dolgert2008}, SQL databases, 
interactive and shareable notebook-based computing \citep{perez2007}. For all 
of these reasons, Python is an ideal platform for a modern data-
analysis environment for solar physics.

SunPy is built on top of the core scientific Python packages namely 
\texttt{NumPy}, \texttt{SciPy}, \texttt{matplotlib}. These core packages 
provide capabilities surpassing that provided by IDL for array manipulation and 
2D plotting \citep{greenfield2011}. Providing similar capabilities as SSW/gen 
in Python is made easier by leveraging the diverse Python community. For 
example, SunPy makes use of the \texttt{pandas} package to provide 
high-performance, easy-to-use data structures and data analysis tools for time 
series data\citep{mckinney2012}. \texttt{pandas} was originally developed for 
quantitative analysis of financial data and has since grown into a generalised 
data analysis package. SunPy also relies on and is coordinating with the 
\texttt{Astropy} package \citep{theastropycollaboration2013}. \texttt{Astropy} is still in 
active development but already provides many relevant capabilities such as 
full-featured support for FITS files, absolute date and time conversions, 
units and unit conversions, and model-fitting. Pandas (134k lines of code) 
and Astropy (90k lines of code) allow SunPy to leverage a total value of \$7.6M 
worth of development which is close to the \$10.4M for SSW/gen assuming a rough 
equivalency between a line of Python code and a line of IDL code.

>>>>

\section{Comparison to SolarSoft}\label{sec:ssw}
This section compares the SunPy library that is described in Sections 
\ref{sec:DataTypes}--\ref{sec:util} with the current de-facto standard solar 
physics data analysis library: SolarSoft (SSW) \citep{freeland1998}.
SolarSoft is a set of integrated software libraries, databases, and system 
utilities which provide a common programming and data-analysis environment for 
solar physics. SSW is modular in nature, which allows users to add required 
packages which contain specialised software for different instrument or tools.

SSW relies upon IDL (Interactive Data 
Language), a commercial and closed-source interactive data analysis environment 
sold by Exelis, a global aerospace, defence, information and services company. 
The SSW `gen' package provides the base analysis capabilities such as time-series
analysis, time conversions, spectral fitting, image display, and file 
I/O. It is composed of 89\% IDL code while the remainder is mostly Perl (5.4\%) 
and csh scripts (3.8\%). 
According to data generated using David A. Wheeler's `SLOCCount'\footnote{\url{http://www.dwheeler.com/sloccount/}}, SSW/gen is composed of 
289,724 lines of code with a total approximate value of \$10.4M (based on the 
Basic COCOMO model \citep{_cocomo_2014}). 
%dps - This data was calculated using...} If we can add them to the repository 
%will improve the quality of the paper and we can always repeat the calculation in the future.
%schriste - i don't know what you mean by this 
%Stuart: comment was Davids
Individual missions and instruments provide additional optional packages to 
support and process their data. For example, the SDO package \citep{sdo} is 
100\% IDL code and is 11k lines of code (\$360k) and the SOHO package 
\citep{soho} is 100k lines of code (\$3M) and is 99.9\% IDL code. SSW is hosted 
and distributed by the Lockheed Martin Solar and Astrophysics Laboratory.
%The RHESSI package is 99.97\% IDL code and is 145k total lines of code (\$5M)
While SSW itself is open-source and freely available, the IDL core is not. The cost 
of IDL is prohibitive for many undergraduates as well as institutions 
with limited means.
Even though SSW's source is open, the development of the source is primarily closed,
with no open issue tracker or established mechanism for community contributions.

One of SunPy's key aims is to provide a free and modern alternative to the 
SolarSoft (gen) library. This is made possible by the recent rise of the 
scientific Python community. Python, as described in Section \ref{sec:Intro}, 
is a general purpose high-level programming language which is free to use and 
cross-platform and powerful programming language.
It therefore has a very large user community which extends far outside of the 
solar physics or astronomy communities. It is very extensible with C, C++, 
FORTRAN or even IDL. Many books and on-line documentation is available. For 
example, at the time writing a search 
%
%dps - \footnote{We need to be careful with this! IDL means more things that 
%our 
%beloved Interactive Data Language. If we search for that and programming 
%then we get just 7 books! - from which 5 are just the same. I don't know how we 
%could make a more accurate search.
%If we script it it would be amazing!  Also, I would just provide the order
%of magnitude, and not a exact number, i.e. 190 and 1800 results}
% 
% schriste - I think we defined IDL properly. I was being generous with the 
%book search for IDL
%Stuart: again not not mine, I think we should drop the book thing, in 
%comparison to the other metrics used it's a little weak.
on Amazon for books on IDL programming generated $189$ results while a search 
for Python yielded $1\,788$ results. Python has a strong emphasis on 
readability a concept which is well-aligned with the scientific endeavour.
Additionally, Python is increasingly being taught in universities, which means 
that research projects can begin quickly without the need for students to learn 
a new (and relatively niche) language. Finally, knowledge of Python is a skill 
which is useful and well compensated outside of the scientific community 
(average salary of \$110k in the Washington, DC, area according to indeed.com at 
the time of writing). Python continues to see increased use in the astronomy 
community \citep{greenfield2011} which has similar goals and requirements as 
the solar physics community. Finally, Python integrates well with many 
technologies such as web servers \citep{dolgert2008}, SQL databases, 
interactive and shareable notebook-based computing \citep{perez2007}. For all 
of these reasons, Python is an ideal platform for a modern data-
analysis environment for solar physics.

SunPy is built on top of the core scientific Python packages namely 
\texttt{NumPy}, \texttt{SciPy}, \texttt{matplotlib}. These core packages 
provide capabilities surpassing that provided by IDL for array manipulation and 
2D plotting \citep{greenfield2011}. Providing similar capabilities as SSW/gen 
in Python is made easier by leveraging the diverse Python community. For 
example, SunPy makes use of the \texttt{pandas} package to provide 
high-performance, easy-to-use data structures and data analysis tools for time 
series data \citep{mckinney2012}. \texttt{pandas} was originally developed for 
quantitative analysis of financial data and has since grown into a generalised 
data analysis package. SunPy also relies on and is coordinating with the 
\texttt{Astropy} package \citep{theastropycollaboration2013}. \texttt{Astropy} is still in 
active development but already provides many relevant capabilities such as 
full-featured support for FITS files, absolute date and time conversions, 
units and unit conversions, and model-fitting. Pandas (134k lines of code) 
and Astropy (90k lines of code) allow SunPy to leverage a total value of \$7.6M 
worth of development which is close to the \$10.4M for SSW/gen assuming a rough 
equivalency between a line of Python code and a line of IDL code.
